---
title: "EliteProspects API Cookbook"
author: "Philip Bulsink"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

#Introduction
This vignette is based on the [EliteProspects API Cookbook](https://github.com/menmo/eliteprospects-api-documentation/blob/master/cookbook.md) by [@peterstark72](https://github.com/peterstark72). It will serve as a set of examples for using the EliteProspectsR package.

```{r library}
library(EliteProspectsR)
```

#Getting A Player's Career Stats

You want the complete Wayne Gretzky career stats.

In order to get the career stats for a specific player, you need the player's EPID (EliteProspects ID). The easiest way to get the EPID is by searching for the player at [eliteprospects.com](eliteprospects.com) and extract the EPID form the URL of the player page. Wayne Gretzky's page is [http://www.eliteprospects.com/player.php?player=9678](http://www.eliteprospects.com/player.php?player=9678). The EPID is the last digits (9678).

##API Call
With the EPID, we can query the API:

```{r playerstats}
gretzky <- getPlayerStats(player_id = 9678)
```

##Notes

The query returns a list of PlayerStats objects. Each object represents the player's stats for a specific League, Season, GameType and Team. For exampe, if you just want Wayne Grezky's NHL playoff stats, `filter` by `list("gameType"="PLAYOFFS", "league.id"="7")` (since the NHL league ID is 7). More information can be found in the Filter & Sorting vignette.

#Getting EPID for an NHL player by name

Maybe you want to get the EPID for an NHL player so that you can access the player's complete EP profile, without having to go online and retrieve the EPID manually as we did above. .

##API Call
Lets say you want the EPID for Sidney Crosby. Use the search function with "q=Sidney Crosby" and "type=player" to restrict the search to only players (not teams).

```{r search_player}
players_search<-search_EP(query = "Sidney Crosby", type='player')
```

##Notes

Sidney Crosby is a simple example, since you only get one hit. If, instead, we try "Corey Perry" we'll get >50 results because "Perry" is a common name. In this case, since we know Corey Perry is born in 1985, we can restrict the query to players born that year.

```{r search_player_filter}
url<-"http://api.eliteprospects.com/beta/search?type=player&q=Corey Perry&filter=latestPlayerStats.league.id=7&yearOfBirth=1985"
```

Again, see the Filter & Sort vignette for more information on slicing queries.

#Getting Latest Transfers

You want to access the latest transfers as seen on the main page of eliteprospect.com.

##API Call

Query the transfers endpoint with "transferProbability=CONFIRMED" and "transferType=TRANSFER" filters, sorted by "transferDate:desc" and "id:desc":

```{r transfers}
transfers<-listTransfers(filters = list("transferProbability"="CONFIRMED", "transferType"="TRANSFER"), sorts = list("transferDate"="desc", "id"="desc"))
```

The response includes the player object, to- and from-team objects and transfer date.

##Notes

Transfers are created by the EP Staff and set to "CONFIRMED" or "RUMORED" and a transfer date. A new transfer automatically gets a higher "id" than the previous, so by sorting by "id" you get transfers in creation order. There is no straightforward way of accessing transfers created in the last 24h.

There are different types of transfers, defined by the "transferType" property: "TRANSFER", "LOAN", "CONTRAXT_EXTENSION", and "RETURNS".

